### 2021-03-18 ch/2-related-work.tex
Add two references

In a follow up work, \citet{zafar2017fairnesstreatment} adapt their work to enforce \acf{EOdds},
as defined in \citet{hardt2016equality}.
(However, it is here referred to as ``Disparate Mistreatment'' instead of ``equalised odds''.)
As in the previous work, they define a tractable proxy for this measure.
To this end, they first define a function \(g\) that measures
how well the class label \(y \in \{0, 1\}\) and the predicted score for a data point \(0 \leq f(x) \leq 1\) agree:
\begin{align}
  \label{eq:zafar-constraint-2}
  g(y, x) = \min \left(0, \left(y - \tfrac{1}{2}\right) \cdot \left(f(x) - \tfrac{1}{2}\right)\right)
\end{align}
Subtracting \(\tfrac{1}{2}\) is necessary to centre the values around \(0\).
The proxy for \ac{EOdds} is then the covariance between the sensitive attribute and \(g\).
\(g\) can also be modified to only take into account cases with \(y=1\) which corresponds to Equality of Opportunity.
The paper contains experiments on the ProPublica/COMPAS dataset.
The task for this dataset is to predict
whether a criminal offender committed another misconduct or felony within two years,
based on personal information that includes age, gender, race and past criminal history.
% The authors compare the proposed method to an implementation of the algorithm from \citet{hardt2016equality}.
% In the Equality of Opportunity task, the approach by \citet{hardt2016equality} performs better
% (lower discrimination, similar accuracy).
% The better performance is likely at least partially explained
% by \citet{hardt2016equality} using the sensitive attribute for predictions.
% The evaluation metric is
% the difference in FPR (false positive rate) and the difference in FNR (false negative rate) between the two groups.
% Hardt et al.~achieve an FPR difference of 0.02 at 66\% accuracy while Zafar et al.~achieve an FPR difference of
% 0.06 at the same 66\% accuracy.
% For Equalised Odds,
% the proposed approach achieves higher accuracy but also higher discrimination than in \citet{hardt2016equality}.
% One limiting factor that the authors identify is the size of the dataset:
% the degree of discrimination cannot be accurately estimated by the algorithm if the dataset is small.

@@

\section{Groundtruth-centric view of bias}%
\label{sec:groundtruth-centric-view-of-bias}
The fairness-accuracy trade-off which results from fairness constraints like \ac{DP} and \ac{EOpp}
has often been seen as a thorny issue.
\citet{wick2019unlocking} were one of the first to argue for a change of perspective to solve this.
For example, the labels of the ProPublica/COMPAS dataset had been treated as unbiased ground truth in previous work,
which the authors call into question.
The simple change they do is evaluating fairness-enforcing models on actual \emph{unbiased} test data.
In practice, they simulate datasets such that unbiased labels can be used for evaluation,
which is necessary because they do not have access to (unobserved) ground truth labels.
The paper considers label bias and ``selection bias''
(``selecting a subsample of the data in such a way that happens to introduce unexpected correlations,
say, between a protected attribute and the target label''),
which we previously referred to as sample bias.
In their experiments they find,
that enforcing fairness can in some situations improve the accuracy on the (unbiased) test set.

>>>>

In a follow up work, \citet{zafar2017fairnesstreatment} adapt their work to enforce \acf{EOdds},
as defined in \citet{hardt2016equality}.
(However, it is here referred to as ``Disparate Mistreatment'' instead of ``equalised odds''.)
As in the previous work, they define a tractable proxy for this measure.
To this end, they first define a function \(g\) that measures
how well the class label \(y \in \{0, 1\}\) and the predicted score for a data point \(0 \leq f(x) \leq 1\) agree:
\begin{align}
  \label{eq:zafar-constraint-2}
  g(y, x) = \min \left(0, \left(y - \tfrac{1}{2}\right) \cdot \left(f(x) - \tfrac{1}{2}\right)\right)
\end{align}
Subtracting \(\tfrac{1}{2}\) is necessary to centre the values around \(0\).
The proxy for \ac{EOdds} is then the covariance between the sensitive attribute and \(g\).
\(g\) can also be modified to only take into account cases with \(y=1\) which corresponds to Equality of Opportunity.
The paper contains experiments on the ProPublica/COMPAS dataset~\citep{angwin2016machine}.
The task for this dataset is to predict
whether a criminal offender committed another misconduct or felony within two years,
based on personal information that includes age, gender, race and past criminal history.
% The authors compare the proposed method to an implementation of the algorithm from \citet{hardt2016equality}.
% In the Equality of Opportunity task, the approach by \citet{hardt2016equality} performs better
% (lower discrimination, similar accuracy).
% The better performance is likely at least partially explained
% by \citet{hardt2016equality} using the sensitive attribute for predictions.
% The evaluation metric is
% the difference in FPR (false positive rate) and the difference in FNR (false negative rate) between the two groups.
% Hardt et al.~achieve an FPR difference of 0.02 at 66\% accuracy while Zafar et al.~achieve an FPR difference of
% 0.06 at the same 66\% accuracy.
% For Equalised Odds,
% the proposed approach achieves higher accuracy but also higher discrimination than in \citet{hardt2016equality}.
% One limiting factor that the authors identify is the size of the dataset:
% the degree of discrimination cannot be accurately estimated by the algorithm if the dataset is small.

@@

\section{Groundtruth-centric view of bias}%
\label{sec:groundtruth-centric-view-of-bias}
The fairness-accuracy trade-off which results from fairness constraints like \ac{DP} and \ac{EOpp}
has often been seen as a thorny issue.
\citet{wick2019unlocking} were one of the first to argue for a change of perspective to solve this.
For example, the labels of the ProPublica/COMPAS~\citep{angwin2016machine} dataset had been treated as unbiased ground truth in previous work,
which the authors call into question.
The simple change they do is evaluating fairness-enforcing models on actual \emph{unbiased} test data.
In practice, they simulate datasets such that unbiased labels can be used for evaluation,
which is necessary because they do not have access to (unobserved) ground truth labels.
The paper considers label bias and ``selection bias''
(``selecting a subsample of the data in such a way that happens to introduce unexpected correlations,
say, between a protected attribute and the target label''),
which we previously referred to as sample bias.
In their experiments they find,
that enforcing fairness can in some situations improve the accuracy on the (unbiased) test set.
